---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(tidyverse)
library(nycflights13)
options(scipen = 999)
```

# Tutorial - Mutate & Summarise

## Mutate

`mutate()` adds new variables and preserves existing ones; transmute() adds new variables and drops existing ones. New variables overwrite existing variables of the same name. Variables can be removed by setting their value to NULL.

The mutate function must return a vector of the same length as the tibble/dataframe used in the data argument, or for grouped data, the length of the group. Functions that are called inside mutate can take column names as an argument which will input the column **as a vector** (or sections of the vector by group).

### Anatomy of a mutate function

Basic mutate calls have a simple construction. Each new column needs a name, each column is assigned an atomic vector (a vector of the same data type).

```{r eval=FALSE, include=TRUE}

mutate(data,{column name} = {vector})

data %>%
   mutate(column1 = function(val1,val2))

# is equivalent to
data %>%
   mutate(column1 = function(data$val1,data$val2))
```

## Evaluation

Below are some examples of mutate 

```{r}
print_and_return <- function(x) {
  print(x)
  return(x)
}

example_data <- tibble(A = 1:4,grp = rep(LETTERS[1:2],each = 2))

example_data %>% 
  mutate(res = print_and_return(A))

# Vectors by group
example_data %>%
  group_by(grp) %>% 
  mutate(res = print_and_return(A))

# Rowwise function
example_data %>%
  rowwise() %>% 
  mutate(res = print_and_return(A))
```

### Vectorised functions

```{r}
tibble(A = 1:3,B = 4:6) %>% 
  mutate(C = sum(A,B),
         D = A + B)

tibble(A = 1:3,B = 4:6) %>% 
  mutate(C = max(A,B),
         D = pmax(A, B))
```

Vectorisation is faster. Friends don't let friends use rowwise.

```{r}
dat <- tibble(A = rnorm(1E5),B = rnorm(1E5))
start_time = Sys.time()
dat %>% 
  rowwise() %>% 
  mutate(C = sum(A,B)) %>% 
  print(n = 3)

Sys.time() - start_time

start_time = Sys.time()
dat %>% 
  mutate(D = A + B) %>% 
  print(n = 3)
Sys.time() - start_time
```

If a single result is given for a group, it is repeated for the length of the group.

```{r}
tibble(A = rnorm(5)) %>% 
  mutate(B = 1,
         C = mean(A))
```

### Common use cases

#### Calculating percentages of a single row as part of a group

```{r warning=FALSE}
planes %>% 
  group_by(type,engines) %>% 
  summarise(n = n(),.groups = "drop_last") %>% 
  # last group in group by is dropped
  # This call takes
  mutate(sum_n = sum(n),
         pct = n/sum_n,
         pct_label = scales::percent(pct))
```

## Summarise

`summarise()` creates a new data frame. It will have one (or more) rows for each combination of grouping variables; if there are no grouping variables, the output will have a single row summarising all observations in the input. It will contain one column for each grouping variable and one column for each of the summary statistics that you have specified.
