---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(tidyverse)
```

# Tidyverse workhorse functions

## The functions you will use all the time

-   %>%

-   select

-   filter

-   group_by

-   mutate

-   summarise

-   head/tail/slice

-   arrange

-   tibble

## Pipes - %>%

Pipes an object forward into a function or call expression. By defauly, pipes take the result of the left hand side expression and puts it in the first argument of the following function/call, without having to specify the argument. It allows you to string together functions step-by-step, without saving the intermediate steps to variables. Tidyverse functions typically set their first argument as the *data* argument, ww . The short-cut for entering a pipe in RStudio is cntrl + shift + M.

```{r}
select(mtcars,vs)

mtcars %>% 
  select(vs)

mtcars %>% 
  rownames_to_column(var = "carName") %>% 
  mutate(make = str_extract(carName,"\\w+")) %>% 
  group_by(make) %>% 
  summarise(mpg = mean(mpg)) %>% 
  arrange(-mpg) %>% 
  head(10)


mtcars %>% {select(.,vs)}
```


## select

```{r}
# Reference by individual name
mtcars %>% 
  select(mpg,cyl)

# Refernce by index
mtcars %>% 
  select(1,2)

# Select a range
mtcars %>% 
  select(mpg:hp,6:8)

# Deselect
mtcars %>% 
  select(-c(mpg,hp))

# Select by string - all
col_names_all <- c("mpg","cyl")
mtcars %>% 
  select(all_of(col_names_all))

# Select by string - all
col_names_any <- c("mpg","cyl","duck")
mtcars %>% 
  select(any_of(col_names_any))

# Error
col_names_all <- c("mpg","cyl")
mtcars %>% 
  select(all_of(col_names_all))

# Reorder, include everything
mtcars %>% 
  select(vs,am,everything())

# Rename columns during select, note it will only keep those listed
mtcars %>% 
  select(cylinders = cyl)
```

## filter

```{r}
flights <- nycflights13::flights

# Numerical filtering
flights %>%
  filter(month >= 6, day == 1)

# is eqivalent to
flights %>%
  filter(month >= 6 & day == 1)

# Filter for specific values
flights %>%
  filter(month %in% c(1,3,5,7,9))

# Betwween two numbers
# Between matches EQUAL and larger/smaller. 
flights %>% 
  filter(between(month,6,8)) %>% 
  count(month)

# Grouped filters
flights %>% 
  group_by(tailnum) %>% 
  filter(all(dep_delay < 0))

# String filters
nycflights13::airports %>% 
  filter(name == "Williams County Airport")

nycflights13::airports %>% 
  filter(str_detect(name,"County"))
```


##  mutate

```{r}
flights %>% 
  mutate(ave_speed = distance/air_time*60)

flights %>% 
  transmute(arr_delay,
            on_time1 = arr_delay < 0,
            on_time2 = if_else(arr_delay > 0,"Late","On-time"),
            # Functions that return a single result will be repeated in each row
            ave_delay_mean = mean(arr_delay,na.rm = T),
            ave_delay = arr_delay > 0)

flights %>% 
  transmute(across(c(dep_time,arr_time),as.character),
            dep_time = str_pad(dep_time,4,pad = "0",side = "left"),
            dep_hour = str_sub(dep_time,start = 1,end = 2),
            dep_min = str_sub(dep_time,start = 3,end = 4))
```


##  summarise

```{r}
flights %>% 
  summarise(n = n(),
            mean_late = mean(arr_delay,na.rm = T),
            sd = mean(arr_delay,na.rm = T))


flights %>% 
  group_by(origin) %>% 
  summarise(n = n(),
            mean_late = mean(arr_delay,na.rm = T),
            sd = mean(arr_delay,na.rm = T))

flights %>% 
  group_by(origin) %>% 
  summarise(unique_destinations = n_distinct(dest),
            unique_carriers = n_distinct(carrier))

```


##  head/tail/slice

```{r}
flights %>% 
  head()

flights %>% 
  tail(10)

flights %>% 
  group_by(tailnum) %>% 
  arrange(month,day,dep_time) %>% 
  slice(1)

flights %>% 
  group_by(tailnum) %>% 
  arrange(month,day,dep_time) %>% 
  slice(1)

flights %>% 
  arrange(distance) %>% 
  slice(1:5,(n() - 4):n())

flights %>% 
  group_by(tailnum) %>% 
  slice_max(distance,n = 1,with_ties = F)
```


## group_by
```{r}

print_return <- function(x) {
  print(x)
  x
}

example_data <- 
  tibble(dat = 1:6,grp = rep(c("A","B"),times = 3)) 

example_data %>% 
  mutate(ret = print_return(dat),
         min_dat = min(ret))

example_data %>% 
  group_by(grp) %>% 
  mutate(ret = print_return(dat),
         min_dat = min(ret))  
```


##  arrange

```{r}
flights %>% 
  distinct(carrier,month,day) %>% 
  arrange(desc(carrier),-month,-day) %>% 
  select(carrier,month,day) 
  
```

