[aoc day 1](https://adventofcode.com/2021/day/2)

Task: Calculate the horizontal position and depth you would have after following the planned course. What do you get if you multiply your final horizontal position by your final depth?

-   forward X increases the horizontal position by X units
-   down X increases the depth by X units
-   up X decreases the depth by X units

<https://adventofcode.com/2021/day/1>

```{r}
library(here)
library(tidyverse)
library(magrittr)
data_file <- here("data/aoc_data/aoc_day2.txt")
aoc_day2 <- readr::read_lines(data_file)
test <- readr::read_lines("forward 5
down 5
forward 8
up 3
down 8
forward 2")
```

## Part 1

```{r}
results <- tibble(reading = aoc_day2) %>% 
  mutate(direction = str_extract(reading,"\\w+"),
         distance = str_extract(reading,"\\d+")) 
  
results %>% 
  count(direction)

results %>% 
  count(distance)

results %>% 
  mutate(plane = if_else(direction == "forward","horizontal","vertical"),
         distance = as.integer(distance),
         distance = if_else(direction == "up",-distance,distance)) %>% 
  group_by(plane) %>% 
  summarise(position = sum(distance))
```

## Part 2

In addition to horizontal position and depth, you'll also need to track a third value, aim, which also starts at 0. The commands also mean something entirely different than you first thought:

-   down X increases your aim by X units.
-   up X decreases your aim by X units.
-   forward X does two things:
-       It increases your horizontal position by X units.
-       It increases your depth by your aim multiplied by X.

```{r}
results %>% 
  mutate(plane = if_else(direction == "forward","horizontal","vertical"),
         distance = as.integer(distance),
         distance = if_else(direction == "up",-distance,distance),
         aim = cumsum(if_else(plane == "vertical",distance,0L)),
         horiz_pos = cumsum(if_else(plane == "horizontal",distance,0L)),
         depth = if_else(plane == "horizontal",aim * distance,0L),
         depth_cumsum = cumsum(depth)) %>% 
  tail(1) %$%
  (horiz_pos * depth_cumsum)
```
