[aoc day 3](https://adventofcode.com/2021/day/3)

Task: 



```{r}
library(here)
library(tidyverse)

data_file <- here("data/aoc_data/aoc_day3.txt")
aoc_day3 <- readr::read_lines(data_file)
```

## Part 1

```{r}
#results <- 
  tibble(reading = aoc_day3)
  
  
```

## Part 2

In addition to horizontal position and depth, you'll also need to track a third value, aim, which also starts at 0. The commands also mean something entirely different than you first thought:

-   down X increases your aim by X units.
-   up X decreases your aim by X units.
-   forward X does two things:
-       It increases your horizontal position by X units.
-       It increases your depth by your aim multiplied by X.

```{r}
results %>% 
  mutate(plane = if_else(direction == "forward","horizontal","vertical"),
         distance = as.integer(distance),
         distance = if_else(direction == "up",-distance,distance),
         aim = cumsum(if_else(plane == "vertical",distance,0L)),
         horiz_pos = cumsum(if_else(plane == "horizontal",distance,0L)),
         depth = if_else(plane == "horizontal",aim * distance,0L),
         depth_cumsum = cumsum(depth)) %>% 
  tail(1) %$%
  (horiz_pos * depth_cumsum)
```
