---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r}
library(tidyverse)

diamonds
```

1.  Explore the distribution of each of the x, y, and z variables in diamonds. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.

```{r}
diamonds %>% 
  ggplot(aes(x)) +
  geom_density()

diamonds %>% 
  select(x,y,z) %>% 
  pivot_longer(everything(),names_to = "name",values_to = "value") %>% 
  ggplot(aes(value,name,fill = name)) +
  geom_boxplot()
```


2.  Explore the distribution of price. Do you discover anything unusual or surprising? (Hint: Carefully think about the binwidth and make sure you try a wide range of values.)

```{r}
diamonds %>% 
  ggplot(aes(price)) +
  geom_histogram(binwidth = 50)
```

```{r}
diamonds %>% 
  mutate(cuts = cut_interval(price,length = 10)) %>% 
  filter(between(price,500, 5000)) %>% 
  count(cuts) %>% 
  arrange(n)
```


3.  How many diamonds are 0.99 carat? How many are 1 carat? What do you think is the cause of the difference?

```{r}
diamonds %>% 
  ggplot(aes(carat)) +
  geom_histogram(binwidth = 0.05) +
  scale_x_continuous(minor_breaks = seq(0,5,0.25))
```


4.  Compare and contrast coord_cartesian() vs xlim() or ylim() when zooming in on a histogram. What happens if you leave binwidth unset? What happens if you try and zoom so only half a bar shows?
```{r}
diamonds %>% 
  ggplot(aes(carat,price)) +
  geom_point(alpha = 0.3)

diamonds %>% 
  ggplot(aes(carat,price)) +
  geom_point(data = ~select(.,-cut),col = "grey",alpha = 0.2) +
  geom_point(aes(col = cut),alpha = 0.5) +
  facet_wrap(~cut)

```

